---
title: "queryAPI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{queryAPI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
<img src=https://www.inaturalist.org/attachments/project_assets/827-VCE_atlas_top-sunrisemtns-990x400.jpg?1551451641>

Download the package and load the library

```{r, eval = FALSE}
# this only needs to be run if:
# 1) downloading for the first time 
# 2) updating to get new functions

devtools::install_github("VtEcostudies/VTatlas", ref = "main")
```

```{r setup}
library(VTatlas)
```

# Using the <code>queryAPI</code> function

The <code>queryAPI</code> function grabs occurrence records using the Vermont Atlas of Life's API. 
See VAL's API guide <a href=https://biocache-ws.vtatlasoflife.org/oldapi blank = true>here</a>

## Simple query 

Below is a simple query where we return all the records that have 'Odonata' in the order field.
by default the function returns all the records. One could limit the amount of 
records that are returned by setting the <code>pageSize</code> argument. If a 
<code>pageSize</code> set by the user is larger than the number of observations
the data frame returned will be filled with NAs.

```{r}
Odonata <- queryAPI(query = list('order:Odonata'))

head(Odonata[,1:5])
```

Manually set pageSize.

```{r}
Odonata_1000 <- queryAPI(query = list('order:Odonata', pageSize = 1000))

tail(Odonata_1000[,1:5])
```

Spatial query

```{r}
data(towns)# <- sf::st_read("data-raw/VTtownsWGS84.shp")

wkt_text <- sf::st_as_text(sf::st_geometry(towns[towns$TOWNNAME=="BENNINGTON",]))
```

```{r}
Od_BEN <- queryAPI(query = list('order:Odonata'),
                   spatialWKT = wkt_text)
```

```{r, echo = FALSE, fig.width = 5, fig.height = 8}
par(mar = c(0,0,0,0))
plot(sf::st_geometry(towns), border = "gray")
plot(sf::st_geometry(towns[towns$TOWNNAME=="BENNINGTON",]),
     add = TRUE, col = "gray")
points(Od_BEN$decimalLatitude~Od_BEN$decimalLongitude,
       pch = 20, col = "red")
points(Odonata$decimalLatitude~Odonata$decimalLongitude, pch = ".")
legend('bottomright', 
       legend = c("Odonata observations", "Bennington observations"),
       pch = c(20,20),
       col = c('black','red'), 
       bty = "n")
```
